Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)

### 2016-03-09 manuscript/supplementary.Rnw
More edits to text. Include updated priors for mortality models and changes to captions

The structure and parameter definitions differ slightly to that in equation \ref{eq:r_bg}. Here, the model has two additional parameters, $\beta_2$ and $\beta_3$. The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius. The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm). The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively. Note that growth constant, $R$, now varies through time as a function of inter-tussock gap radius.

@@

As with growth rates, we modelled each individual's cumulative hazard per census, $Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches to those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the length (in years) between each census, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.

@@

Like the growth models, the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from weakly informative positive half Cauchy priors with scale parameters of 25. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with a scale parameter of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These Cauchy priors are the recommended default choice for logistic regression models\cite{Gelman:2008ua}.

@@

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census t. However, because some cardinal distances were recorded as censored data (i.e. \textgreater 25 cm), these data were incomplete. We estimated these censored data using a censored model (see below). Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean. Furthermore, standardizing by two standard deviations allows the magnitude of effects to be directly compared between binary and continuous covariates\cite{Gelman:2007te}.

\subsection{\textit{Modelling censored Poa distances.}}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.

@@

This functional form allows gap radius to either increase or decrease exponentially. We did not force this curve to be a negative exponential (i.e. gaps decreased over time) because it is theoretically possible for inter-tussock gaps to increase over time. Here $\alpha_i$ is a known constant that represents the initial inter-tussock gap radius for each gap $i$ at the beginning (i.e. the intercept). $\beta_i$ represents the exponential rate for each gap and $Days_t$ is the number of days from the beginning of the experiment.

@@

Half Cauchy priors with a scale of 25 were used for all standard deviation parameters and the overdispersion parameter $\phi$. Hyperparameter $\mu_site$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 seedlings/m$^2$.

@@

\begin{equation}\label{eq:maxht_obs} 
maxht_i \sim LN(\widehat{\mu}_i, \sigma_{obs}),
\end{equation}

@@

where the intercept, $\alpha_j$, varies by site and represents the average log maximum seedling height of each site, assuming all other parameters are at their mean. We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average log maximum height for the average site and a standard deviation of $sigma_{site}$ which defines site variability. The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and topographic wetness, respectively and $\beta_3$ is the effect of fire severity.

@@

\subsection{\textit{Data inputs into density and maximum height models.}}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity $Severity_j$, was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were r \textless 0.16. The data from each continuous covariate was then centered on zero and standardized by two standard deviations\cite{Gelman:2007te}. 
\pagebreak
\section*{Supplementary figures}
\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}) \& topographic wetness (\textbf{e}), in areas burnt by the 2003 fires. All errors bars are 95\% Bayesian Credible Intervals. Phi = overdispersion parameter for negative binomial distribution.}
\label{figs1}
\end{minipage}
\end{suppfigure}
\pagebreak

@@

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS3.pdf}
\caption{\textbf{Observed and projected diameter growth trajectories of four dominant Australian alpine shrubs.} Thick lines represent mean ($\pm$ 95\% CI) observed heights at each May census. Thin lines represent individual growth trajectories. Red or orange lines = seedlings growing in warmed (OTC) conditions and blue or light blue lines = seedlings growing in control conditions.} 
\label{figs3}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS4.pdf}
\caption{\textbf{Mean maximum seedling height along fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. Centered and standardized model coefficients (on the log scale) (\textbf{a} \& \textbf{b}) and effects of fire severity  (\textbf{c} \& \textbf{d}), altitude (\textbf{e} \& \textbf{f}) \& topographic wetness (\textbf{g} \& \textbf{h}), in areas burnt by the 2003 fires. All errors bars are 95\% Bayesian Credible Intervals.}
\label{figs4}
\end{minipage}
\end{suppfigure}
\pagebreak

>>>>

The structure and parameter definitions differ slightly to that in equation \ref{eq:r_bg}. Here, the model has two additional parameters, $\beta_2$ and $\beta_3$. The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius. The inclusion of these parameters, coupled with $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm). The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively. Note that growth constant, $R$, now varies through time as a function of inter-tussock gap radius.

@@

As with growth rates, we modelled each individual's cumulative hazard per census, \\
$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches to those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the length (in years) between each census, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.

@@

Here the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with scale parameters of 2.5. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with a scale parameter of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These Cauchy priors are the recommended default choice for logistic regression models\cite{Gelman:2008ua}.

@@

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some cardinal distances were recorded as censored data (i.e. \textgreater 25 cm), these data were incomplete. We estimated these censored data using a censored model (see below). Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean. Furthermore, standardizing by two standard deviations allows the magnitude of effects to be directly compared between binary and continuous covariates\cite{Gelman:2007te}.

\subsection{\textit{Modelling censored Poa distances.}}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. $\geq$ 26 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.

@@

This functional form allows gap radius to either increase or decrease exponentially. We did not force this curve to be a negative exponential (i.e. gaps decreased over time) because it is theoretically possible for inter-tussock gaps to increase over time. Here $\alpha_i$ is a known constant that represents the initial inter-tussock gap radius for each gap $i$ (i.e. the intercept). $\beta_i$ represents the exponential rate for each gap and $Days_t$ is the number of days from the beginning of the experiment.

@@

Half Cauchy priors with a scale of 25 were used for all standard deviation parameters and the overdispersion parameter $\phi$. Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 seedlings/m$^2$.

@@

\begin{equation}\label{eq:maxht_obs} 
maxht_i \sim LN(\widehat{maxht}_i, \sigma_{obs}),
\end{equation}

@@

where the intercept, $\alpha_j$, varies by site and represents the average log maximum seedling height of each site, assuming all other parameters are at their mean. We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average log maximum height for the average site and a standard deviation of $\sigma_{site}$ which defines site variability. The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and topographic wetness, respectively and $\beta_3$ is the effect of fire severity.

@@

\subsection{\textit{Data inputs into density and maximum height models.}}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity $Severity_j$, was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were r \textless 0.16. The data from each continuous covariate was then centered on zero and standardized by two standard deviations\cite{Gelman:2007te}. 
\pagebreak
\section*{Supplementary figures}
\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI).} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}) \& topographic wetness (\textbf{e}), in areas burnt by the 2003 fires. All errors bars are 95\% Bayesian Credible Intervals. Phi = overdispersion parameter for negative binomial distribution.}
\label{figs1}
\end{minipage}
\end{suppfigure}
\pagebreak

@@

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS3.pdf}
\caption{\textbf{Observed and projected diameter growth trajectories of four dominant Australian alpine shrubs.} Thick lines represent mean ($\pm$ 95\% CI) observed heights at each May census. Thin lines represent individual growth trajectories. Red \& orange lines = seedlings growing in warmed (OTC) conditions and blue \& light blue lines = seedlings growing in control conditions.} 
\label{figs3}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS4.pdf}
\caption{\textbf{Mean maximum seedling height along fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. Centered and standardized model coefficients (on the log scale) (\textbf{a} \& \textbf{b}) and effects of fire severity  (\textbf{c} \& \textbf{d}), altitude (\textbf{e} \& \textbf{f}) \& topographic wetness (\textbf{g} \& \textbf{h}), in areas burnt by the 2003 fires. All errors are 95\% Bayesian Credible Intervals.}
\label{figs4}
\end{minipage}
\end{suppfigure}
\pagebreak
